---
title: "597_ZhenzhenGe_zg92_Week_07"
author: "Zhenzhen Ge"
date: "11/4/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(curl)
library(readr)
library(tidyr)
library(stringr)
library(choroplethr)
```

## Maternal mortality analysis

- Download the csv file on maternal mortality and store it in a data directory in your project repo

```{r}
# Download the first file into the local git repository as "data_mat.csv"
# Then read in the data to R
data_mat_raw <- read_csv("data_mat.csv")
head(data_mat_raw)
```

- Although the data is tidy, it needs cleaning. Notice that spaces are used instead of commas in numbers. Turn the numbers for maternal mortality inside square brackets into two new variables, lower_bound and upper_bound

```{r, echo=TRUE}
# Clean the data by separate combined information in one cell
# Delete blanks between the numbers
# We have "Year" with value "2012-2013", I decide to separate these rows into two rows
# This is to make sure we can match GNI per capita data by year
data_mat_tidy <- data_mat_raw %>%
  separate(`Maternal mortality ratio (per 100 000 live births)`, into = c("Maternal mortality ratio","bounds"), sep = "\\[") %>%
  separate(bounds, into = c("Lower_bound","Upper_bound"), sep = "-") %>%
  separate(Year, into = c("Year","Year2"), sep = "-") %>%
  gather(key = "Index", value = "Year", 2:3) %>%
  filter(Year != "NA")
data_mat_tidy$`Maternal mortality ratio` <- as.numeric(gsub("[[:space:]]","",data_mat_tidy$`Maternal mortality ratio`))
data_mat_tidy$Lower_bound <- as.numeric(gsub("[[:space:]]","",data_mat_tidy$Lower_bound))
data_mat_tidy$Upper_bound <- as.numeric(gsub("[[:space:]]|\\]","",data_mat_tidy$Upper_bound))
head(data_mat_tidy)
```

- Download the gross national income data
- Tidy this data, so that there is one entry per country per year

```{r, echo=TRUE}
# Download the data into the default directory
data_income_raw <- read_csv("data_income.csv",skip = 1)

# Tidy the data into simple format and clean the spaces
data_income_tidy <- data_income_raw %>%
  gather(key = "Year", value = "GNI_Per_Cap",2:25)
data_income_tidy$GNI_Per_Cap <- as.numeric(gsub("[[:space:]]","",data_income_tidy$GNI_Per_Cap))
head(data_income_tidy)
```

- Merge the two files by country and year, retaining only rows for which you have per capita income and either maternal mortality or percentage attended births

```{r, echo=TRUE}
# Join the two datasets together and change formats of some columns
# Only select columns we will use in future plots
data_join_raw <- left_join(data_mat_tidy,data_income_tidy)
data_join_tidy <- data_join_raw %>%
  filter(is.na(GNI_Per_Cap) == FALSE) %>%
  select(Country, Year, `Maternal mortality ratio`,`Births attended by skilled health personnel (%)`,GNI_Per_Cap)
data_join_tidy$Year <- as.numeric(data_join_tidy$Year)
colnames(data_join_tidy)[3:4] <- c("Maternal_Mortality_Ratio","Birth_Attend_Rate")
head(data_join_tidy)
```

- Make a two polished  and informative graphs, one relating income to maternal mortality and the other relating income to percentage attended births
- Save a csv file with the merged, cleaned, and tidy data, using a suitable name of your choosing in the data folder

```{r, echo=TRUE}
# Now we prepare to plot the relationships between GNI and maternal mortality
# We want to label extremely high or low data points
# We also scrutinize the data and find there's limited data in certain years
# So we need to pick appropriate Year to plot
# For maternal mortality ratio we end up with more data in year 2000 and 1990
# We show the maternal mortality ratio against GNI perCap
data_highMat <- filter(data_join_tidy, Maternal_Mortality_Ratio > 1200)
ggplot(data = filter(data_join_tidy, Maternal_Mortality_Ratio != "NA"), aes(x = GNI_Per_Cap, y = Maternal_Mortality_Ratio)) + geom_point(colour = "grey50") + ggtitle("Maternal Mortality Ratio Per 100,000 Births Against GNI Per Capita") + facet_grid(.~Year)+ ggrepel::geom_label_repel(aes(label = Country), data = data_highMat,alpha = .25)

# The next plot we show the births attended by professionals ratio
# Only a few years have relatively sufficient data
# So we only pick year 2010-2013 
# We also mark a few lowest data points 
data_lowAtt <- filter(data_join_tidy, Year > 2009 & Birth_Attend_Rate < 50)
ggplot(data = filter(data_join_tidy, Birth_Attend_Rate != "NA" & Year > 2009), aes(x = GNI_Per_Cap, y = Birth_Attend_Rate)) + geom_point(color = "grey50") + ggtitle("2010-2013 Births Attended By Skilled Health Personnel (%) Against GNI Per Capita") + facet_wrap(~Year, ncol = 2)+ ggrepel::geom_label_repel(aes(label = Country), data = data_lowAtt,alpha = .25)

# Save the data to default directory
write_csv(data_join_tidy, "data_join_tidy.csv")
```

- Use the country_choropleth() function in the choroplethr package to make a world map of maternal mortality, using the most recent year for each country in the merged dataset you created

```{r, echo=TRUE}
# To plot the country map we need to fine-tune data format
# We choose year 2000 because other years we don't have much data
# We miss a few countries so the map will show some dark regions
data_join_map <- data_join_tidy %>%
  filter(Year == 2000) %>%
  select(Country,Maternal_Mortality_Ratio)
colnames(data_join_map) <- c("region","value")
data_join_map$region <- tolower(data_join_map$region)
country_choropleth(data_join_map, title = "Maternal Mortality Ratio Per 100,000 Births in 2000")
```

